### 2015-10-13 weeklyReport/appendix.tex
10-13-15,problem to return structured code examples

\begin{figure*}[!htb]
 \begin{minipage}{1\textwidth}
\scriptsize 
\begin{tabular}{@{}p{1\textwidth}} 
 \hline

  \multicolumn{1}{c}{Result 7: PaparusCDTEditor} \\ \hline
  \vspace{-4mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
public class PapyrusCDTEditor extends CEditor \{
 protected IAction gmfUndo, gmfRedo;
 protected IAction textUndo, textRedo;
 @Override
 public void createPartControl(Composite parent) \{
  IActionBars actionBars = getEditorSite().getActionBars();
  if((actionBars != null) && (gmfUndo == null)) \{
   gmfUndo =  \uwave{actionBars.getGlobalActionHandler(ITextEditorActionConstants.UNDO)};
    gmfRedo =  \uwave{actionBars.getGlobalActionHandler(ITextEditorActionConstants.REDO)};
  \}
  super.createPartControl(parent);
  if(actionBars != null) \{
   textUndo = actionBars.getGlobalActionHandler(ITextEditorActionConstants.UNDO);
   textRedo = actionBars.getGlobalActionHandler(ITextEditorActionConstants.REDO);
    \uwave{actionBars.setGlobalActionHandler(ITextEditorActionConstants.UNDO, gmfUndo)};
    \uwave{actionBars.setGlobalActionHandler(ITextEditorActionConstants.REDO, gmfRedo)};
   actionBars.updateActionBars();
  \}
  @Override
  public ISourceViewer createSourceViewer(Composite parent, IVerticalRuler ruler, int styles) \{
   final ISourceViewer viewer = super.createSourceViewer(parent, ruler, styles);
   focusListener = new FocusListener() \{
   public void focusLost(FocusEvent e) \{
    if(isDirty()) \{
     syncCpp.syncCDTtoModel();
     Classifier classifier = (Classifier)papyrusTextInstance.getEditedObject();
     doSave(new NullProgressMonitor());
     SyncModelToCDT.syncModelToCDT(classifier);
    \}
     IActionBars actionBars = getEditorSite().getActionBars();
     if(actionBars != null) \{
      if((gmfUndo != null) && (gmfRedo != null)) \{
       actionBars.setGlobalActionHandler(ITextEditorActionConstants.UNDO, gmfUndo);
       actionBars.setGlobalActionHandler(ITextEditorActionConstants.REDO, gmfRedo);
       actionBars.updateActionBars();
     \} \} \}
   public void focusGained(FocusEvent e) \{
    IActionBars actionBars = getEditorSite().getActionBars();
    if(actionBars != null) \{
     if((textUndo != null) && (textRedo != null)) \{
      actionBars.setGlobalActionHandler(ITextEditorActionConstants.UNDO, textUndo);
      actionBars.setGlobalActionHandler(ITextEditorActionConstants.REDO, textRedo);
      actionBars.updateActionBars();
     \} \} \} \};
  return viewer;
\} \}
 \end{Verbatim}
      \vspace{-4mm}
     \\ \hline
       \multicolumn{1}{c}{Result No.9,10: AspectEditorContributor} \\ \hline
  \vspace{-4mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
public class AspectEditorContributor extends MultiPageEditorActionBarContributor \{
 private IEditorPart activeEditorPart;
 private AspectEditor aspectEditor;
 @Override
 public void setActivePage(IEditorPart part) \{
  if (activeEditorPart == part)   return;
  activeEditorPart = part;
  IActionBars actionBars = getActionBars();
  if (activeEditorPart != null && activeEditorPart instanceof ITextEditor) \{
   IActionBars siteActionBars = ((IEditorSite)activeEditorPart.getEditorSite()).getActionBars();
    \uwave{siteActionBars.setGlobalActionHandler(ITextEditorActionConstants.UNDO, getAction((ITextEditor)activeEditorPart, ITextEditorActionConstants.UNDO))};
    \uwave{siteActionBars.setGlobalActionHandler(ITextEditorActionConstants.REDO, getAction((ITextEditor)activeEditorPart, ITextEditorActionConstants.REDO))};
   siteActionBars.updateActionBars();              
  \} else \{
   if (part instanceof AspectEditor) \{  aspectEditor = (AspectEditor) part; \}
   IWorkbenchPartSite site = aspectEditor.getSite();
   if (site instanceof IEditorSite) \{
    ITextEditor textEditor = aspectEditor.getMultipageEditor().getTextEditor();
    IActionBars siteActionBars = ((IEditorSite) site).getActionBars();
    siteActionBars.setGlobalActionHandler(ITextEditorActionConstants.UNDO, getAction(textEditor, ITextEditorActionConstants.UNDO));
    siteActionBars.setGlobalActionHandler(ITextEditorActionConstants.REDO, getAction(textEditor, ITextEditorActionConstants.REDO));
    siteActionBars.updateActionBars();              
 \} \}\}
 protected IAction getAction(ITextEditor editor, String actionID) \{
   return (editor == null ? null : editor.getAction(actionID));
 \} \}
   \end{Verbatim}
      \vspace{-4mm}
     \\ \hline

\end{tabular} 
\caption{Results from SearchCode CSE}
\label{fig:undoEditor}
\end{minipage}
\end{figure*}

\begin{figure*}[!htb]
 \begin{minipage}{1\textwidth}
\scriptsize 
\begin{tabular}{@{}p{1\textwidth}} 
 \hline 
    \multicolumn{1}{c}{Result No 2,3.: AndroidTextEditor} \\ \hline
  \vspace{-4mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
public abstract class AndroidTextEditor extends FormEditor implements IResourceChangeListener \{
 protected void createAndroidPages() \{
  mIsCreatingPage = true;
  createFormPages();
  createTextEditor();
  createUndoRedoActions();
  postCreatePages();
  mIsCreatingPage = false;
 \}
 private void createUndoRedoActions() \{
  IActionBars bars = getEditorSite().getActionBars();
  if (bars != null) \{
    IAction action = mTextEditor.getAction(ActionFactory.UNDO.getId());
    bars.setGlobalActionHandler(ActionFactory.UNDO.getId(), action);
    action = mTextEditor.getAction(ActionFactory.REDO.getId());
    bars.setGlobalActionHandler(ActionFactory.REDO.getId(), action);
    bars.updateActionBars();
  \}\}
 private void createTextEditor() \{
   mTextEditor = new TextEditor();
   int index = addPage(mTextEditor, getEditorInput());
   mTextPageIndex = index;
   setPageText(index, mTextEditor.getTitle());
   IDocumentProvider provider = mTextEditor.getDocumentProvider();
   mDocument = provider.getDocument(getEditorInput());
   mDocument.addDocumentListener(new IDocumentListener() \{
    public void documentChanged(DocumentEvent event) \{
    onDocumentChanged(event);
 \}); \} \}
\end{Verbatim}
\vspace{-4mm}
 \\ \hline
  \multicolumn{1}{c}{Result 4: DocumentUndoManagerImpl } \\ \hline
  \vspace{-4mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
public class DocumentUndoManagerImpl implements DocumentUndoManager \{
 private static class UndoableTextChange extends AbstractOperation \{
 protected long fUndoModificationStamp = Document.UNKNOWN;
 protected long fRedoModificationStamp = Document.UNKNOWN;
 protected void reinitialize() \{
   fUndoModificationStamp = Document.STAMP;
   fRedoModificationStamp = Document.STAMP;
  \}
 protected void undoTextChange() \{
  try \{
     if (fDocumentUndoManager.fDocument instanceof Document)
         ((Document)fDocumentUndoManager.fDocument).replace(fStart, fText.length(), fPreservedText,fUndoModificationStamp);
     else
          fDocumentUndoManager.fDocument.replace(fStart, fText.length(), fPreservedText);
            \} catch (BadLocationException x) \{
   \} \}
 public boolean canUndo() \{
   if (isValid()) \{
    if (fDocumentUndoManager.fDocument instanceof Document) \{
      long docStamp = ((Document)fDocumentUndoManager.fDocument).getModificationStamp();
      boolean canUndo = docStamp == Document.STAMP ||  docStamp >= getRedoModificationStamp();
      return canUndo;
  \}
  return false; 
\}
public IStatus undo() \{
  if (isValid()) \{
    fDocumentUndoManager.fireDocumentUndo(fStart, fPreservedText, fText, null, DocumentUndoEvent.ABOUT_TO_UNDO, false);
    undoTextChange();
    fDocumentUndoManager.resetProcessChangeState();
    fDocumentUndoManager.fireDocumentUndo(fStart, fPreservedText, fText, null);
    return Status.OK_STATUS;
  \}
 return IOperationHistory.OPERATION_INVALID_STATUS;
 \}
 protected void redoTextChange() \{
  try \{
    if (fDocumentUndoManager.fDocument instanceof Document)
       ((Document)fDocumentUndoManager.fDocument).replace(fStart, fEnd - fStart, fText, fRedoModificationStamp);
   else
     fDocumentUndoManager.fDocument.replace(fStart, fEnd - fStart, fText);
   \} catch (BadLocationException x) \{
  \}\}
protected void updateTextChange() \{
   fText = fDocumentUndoManager.fTextBuffer.toString();
   fDocumentUndoManager.fTextBuffer.setLength(0);
   fPreservedText = fDocumentUndoManager.fPreservedTextBuffer.toString();
   fDocumentUndoManager.fPreservedTextBuffer.setLength(0);
\}
 protected UndoableTextChange createCurrent() \{
   if (fDocumentUndoManager.fFoldingIntoCompoundChange)
     return new UndoableCompoundTextChange(fDocumentUndoManager);
  return new UndoableTextChange(fDocumentUndoManager);
 \} \}
  \end{Verbatim}
      \vspace{-4mm}
     \\ \hline
\end{tabular} 
\caption{Results from SearchCode CSE}
\label{fig:undoEditor}
\end{minipage}
\end{figure*}