### 2015-10-13 weeklyReport/eval-table.tex

\begin{table*}[ht]
\begin{center}
\caption{Evaluation dataset for feature location tools and concern mining tools}
\label{tab:total}
\scriptsize{
\begin{tabular*}{1\textwidth}{@{}l|l|l@{}} \hline
Tool&Input: query / context&Output: method / code snippet\\\hline
FCA~\cite{Denys:FCA12}&jEdit: support for `thick' caret &6 (org.eclipse.swt.widgets.Table.createHandle(), createWidget(), etc)\\
&jEdit: rename project source&5 (org.eclipse.core.internal.localstore.FileSystemStore.move, etc)\\ \hline
%FindConcept~\cite{Hill:FindConcept07}&Add Textfield&5\\
%&Compile Report&8\\ \hline
%COMMIT~\cite{Hassan:concernHistory10}&Error logging&It focuses on improving crosscutting concerns using co-evolving history.\\
%&Debugging&\\ \hline
Portfolio~\cite{Portfolio:DenysICSE11}&mip map dithering texture image graphics&6 methods: ImageTexture:createTextureFromImage, TiledTexture:\\ 
Export~\cite{Export:DenysASE13}&&buildGaussianGlareTexture, LoadTextureFromFile:initReaderer, etc  \\
&adjust parameters such as brightness for a picture*\\
&read and play midi files*\\ \hline
Rastkar, Murphy&jEdit: add autosave capabilities&They assume that there already exist multiple instances of this concern.\\
 et al.~\cite{Murphy:nlConcern11}&add undo functionality for a command&`All of the methods named `undo': \\ 
 &&override method `org.jhotdraw.util.UndoableAdapter.undo';\\
 &&are a member of the class `undoActivity';\\ \hline
 Strathcona~\cite{Holmes:structural05}&update status line&protected void updateStatusLine(IStructuredSelection selection) \{//fill \\
&&String msg = getStatusLineMessage(selection);\\
&&getViewSite().getActionBars().getStatusLineManager().setMessage(msg); \}\\ 
&&\\
&generate method signature&public boolean visit(MethodDeclaration node) \{//fill\\
&& node.getModifiers();\\
&& node.getName.getIdentifiers();\\
&& return super.visit(node); \} \\ \hline
 GraPacc~\cite{Nguyen:GraPaccICSE12}&SWT usage pattern&Display display = new Display();\\
 && Shell shell = new Shell(display); // fill the rest\\
 && shell.open()\\
 && while (!shell.isDisposed()) \{\\
 &&  if (!display.readAndDispatch())\\
 &&   display.sleep();\\
 && \}\\
 && display.dispose();\\ \hline

\end{tabular*}
 \label{tab:concern}

* represents tasks given to the users. However, the free-form query users use to finish the task is unknown.
}

 \end{center}
\end{table*}

\begin{table*}[ht]
\begin{center}
\caption{Evaluation dataset for reuse tool }
\label{tab:total}
\vspace{1mm}
\scriptsize{
\begin{tabular*}{1\textwidth}{@{}l|l|l|l@{}} \hline
Tool&Task&LOC&Result / Description\\\hline
Gilligan~\cite{Holmes:reuse07}&reuse BMP image file parsing in SWT&497&They try to address the problem of source code integration \\
&reuse  module that serializes objects to XML&900& in the context of medium or large-scale reuse tasks. They  \\
&reuse virtual file system from a third-party app&3000& suggest program elements that are easy to reuse based on \\
&reuse GraphML parsing code from Jung&200& structural  relevance and cost of reuse in the source  \\ 
&reuse charting component from Azureus&& context,  and guide  users to  plan a non-trivial reuse task.  \\
&reuse NMEA parser from the Gpsylon project&&They assume that developers have a perfect example at \\
&reuse the lines of code counter from Metrics project&&hand,  and they can finish the reuse task by resolving all\\
&reuse QIF parser from jGnash project&& dependency conflicts.\\ \hline

\end{tabular*}
 \label{tab:reuse}

LOC represents the lines of code that are reused.

}
 \end{center}
\end{table*}

\begin{table*}[ht]
\begin{center}
\caption{Evaluation dataset for code search tools that use free-form query or keyword query}
\label{tab:total}
\vspace{1mm}
\scriptsize{
\begin{tabular*}{1\textwidth}{@{}l|l|l@{}} \hline
Tool&Input: query / keyword&Output: code snippet / code element\\\hline
SNIFF~\cite{sniff:Sen09}&get active editor window from&IWorkbenchWindow window = PlatformUI.\\
& eclipse workbench&getWorkbench().getActiveWorkbenchWindow();\\
&parse a java source and create ast&ASTParser parser = ASTParser.newParser(AST.JLS3);\\
&&CompilationUnit cu = (CompilationUnit) parser.createAST(null);\\
&connect to a database using jdbc&Connection conn = DriverManager.getConnection(path)\\ \hline
Sourcer~\cite{Sourcerer:SC14}&Find a XML parser&org.apache.jasper.xmlparser\\
&Find the use of Depth First Search&org.jeat.search.dfs.BasicDepthFirstSearch\\
&Find the use of Newton Raphson Method&org.cdk.NewtonRaphsonMethod\\ \hline
anyCode~\cite{Gvero:ICSE15}&copy file fname to destination&FileUtils.copyFile(new File(fname), new File(destination))\\
&does x begin with y&x.startsWith(y)\\
&get the current year&new Date().getYear()\\ \hline
EVOLIZER~\cite{gall:nlQueryICSE10}&What methods call addChart?&ServletUtilities: registerChartForDeletion(File, HttpSession)\\
&What method accesses the&DefaultAxisEditor:attemptLabelFontSelection()\\
& attribute labelFontField?\\
&What are the classes that extend JPanel?&DefaultAxisEditor \\ \hline
Keivanloo et al.~\cite{spotWork:ICSE14}&\{getOptionValue, CommandLine\}&CommandLine.getValue(Option)\\
&\{ISelection, isEmpty\}&ISelection.isEmpty()\\
&\{WebElement, click\}&WebElement.click() \\\hline
\end{tabular*}
 \label{tab:freeform}
}
 \end{center}
\end{table*}