### 2015-11-02 proposal/graph-time.tex

\begin{figure*}[!htb]
    \begin{minipage}{0.49\textwidth}
     \centering
 \includegraphics[width=1\textwidth]{fig/fixRateIntroduce.jpeg}   
     % \scriptsize{C-I represents the group of methods that add new inconsistent names; 
%       
%C-C represents the group of methods without inconsistent names that remain the same during the observation period.}
  \end{minipage}
  \hfill
     \begin{minipage}{0.49\textwidth}
     \centering
 \includegraphics[width=1\textwidth]{fig/fixRateRemove.jpeg}  
   %    \scriptsize{I-C represents the group of methods that remove inconsistent names; 
%      
%I-I represents the group of inconsistent name methods that remain as inconsistent name methods during the observation period; }
    \end{minipage}

      \begin{minipage}{1\textwidth}
           \centering
           \scriptsize{
 \begin{tabular}{l|rr|rr} 
 {Fix rate change {\bf before} ($<0$) vs.~{\bf after} ($>0$)}&  \multicolumn{2}{c|}{Introduction of inconsistent names} & \multicolumn{2}{c}{Removal of inconsistent names}\\
Inconsistent names are added/deleted at Revision 0& Mann-Whitney&Cliff Delta& Mann-Whitney&Cliff Delta\\ \midrule
C-I: methods that inconsistent names are introduced&$p<0.01$&$-0.35$(m)&$p<0.01$&0.48(l)\\ 
I-C: methods that inconsistent names are removed&$p<0.01$&$-0.39$(m)&$p<0.01$&0.35(m)\\ 
I-I: methods with inconsistent names&$p=0.18$&N/A&$p=0.10$&N/A\\ 
C-C: methods without inconsistent names&$p=0.21$&N/A&$p=0.24$&N/A\\ 
 \end{tabular}

 m:medium;  l:large
 }
    \end{minipage}
       \vspace{-3mm}
     \caption{Fix rate change before and after introducing/removing inconsistent names}
   \label{fig:rateChange}
      \vspace{-3mm}
\end{figure*}