### 2015-11-02 proposal/intro.tex

\section{Introduction}

\paragraph{Problem}

Developers spend a lot of effort understanding programs. Considering that software documentation can be out of date or does not exist at all, developers need code to be self-explanatory. Therefore, developers are encouraged to assign meaningful semantic names to program elements. Prior study has shown that naming matters from two aspects: 1) poor code convention can do harm to code comprehension~\cite{buse:readability, butler:csmr10, boogerd:icsm08}. 2)  semantic information is effective to solve some complicated problems in program analysis such as code search~\cite{Denys:portfolio11,  lopes:usageSimFSE10, RecoDoc:ICSE12} and   feature location~\cite{Denys:TSE12, sniafl:TSE06, Hill:locateConcern07}.

We explore the synergy between semantics   investigate several areas that