### 2015-11-17 weeklyReport/appendix.tex
init 11-17

\input{example-purity}

\input{example-usage-new}

\input{table-RQ2}