### 2015-11-17 weeklyReport/example-usage-old.tex

\begin{figure}[!htb]
\begin{minipage}{0.47\textwidth}
\scriptsize 
\begin{tabular}{p{1\textwidth}}
 \hline 
 \multicolumn{1}{c}{(A) A type-use pattern with a commonly used name} \\ \hline
%\midrule
\vspace*{-4mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
/* org.apache.jackrabbit.core.mem.InMemPersistenceManager */
1. FileSystem blobFS;
2. public void close() \{
3.   if (blobFS != null) \{
4.    try \{
5.      blobFS.close();
6.    \} catch (FileSystemException e) \{
7.    \} 
8.    finally \{
9.     blobFS.releaseLock();
10.   \} \} \}
\end{Verbatim}
\vspace{-4mm}
\\  \hline
%\midrule
\multicolumn{1}{c}{(B) A method using a field with inconsistent name } \\ \hline
\vspace{-3mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
/* org.apache.jackrabbit.core.IRepositoryImpl */
1. FileSystem{\bf \hl{repStore}};
2. public void doShutdown() \{
3.  \uwave{if (repStore != null)} \{
4.   try \{
5.     \uwave{repStore.close()};
6.   \} catch (FileSystemException e) \{
7.        log.error("error while closing file system", e);
8.   \} \} \}
\end{Verbatim} 
\vspace{-5mm}
\\  \hline
\multicolumn{1}{c}{(C) A refactoring on the field} \\ \hline
Date: 2/16/2010

Change Log: repStore is a FS rather than a DataStore, lock does not release when shutdown repFS, rename FS for consistency. 
\vspace{-2mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
/* org.apache.jackrabbit.core.IRepositoryImpl */
1. FileSystem\color{blue}{\underline{\underline{\bf repFS}}};
2. public void doShutdown() \{
3.  \color{blue}{\uwave{if (repFS != null)}} \{
4.  try \{
5.   \color{blue}{\uwave{repFS.close()}};
6.  \} catch (FileSystemException e) \{
7.    log.error("error while closing file system", e);
8. \color{blue}{\bf \} finally \{ }
9.      \color{blue}{\bf \uwave{repFS.releaseLock()};}
10.  \} \} \}
\end{Verbatim} 
\vspace{-5mm}
\\  \hline
\end{tabular}
\caption{Inconsistent field identified by type-use pattern analysis}
\label{fig:usageExample}
\end{minipage}
\end{figure}

%%%%%%%Usage pattern example %%%%%%%
 \begin{figure}[!htb]
 \begin{minipage}{0.47\textwidth}
 \scriptsize 
 \begin{tabular}{p{1\textwidth}}\hline
   \multicolumn{1}{c}{(A) A type-use pattern with a commonly used name} \\ \hline
 %\midrule
 \vspace{-4mm}
 \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
 /* org.apache.hadoop.hbase.regionserver.LogRoller */
1. FlushRequester requester = null; ...
2. private void scheduleFlush(final byte[] region) \{...
3.   if (requester != null) \{
4.      requester.request(this);...
5.   \}
6. \} 
 \end{Verbatim}
 \vspace{-4mm}
\\ \hline
  \multicolumn{1}{c}{(B) A method with an inconsistent parameter and its caller} \\ \hline
 %\midrule
 \vspace{-4mm}
 \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
 /* org.apache.hadoop.hbase.regionserver.HRegion */
1. Context c = getContext();
2. public static HRegion newHRegion(...,
3.     FlushRequester \hl{flushListener}) \{  
4.   if (\uwave{flushListener} != null)  
5.     \uwave{flushListener}.request(this);
6.   return c.newInstance(..., \uwave{flushListener});  
7.\}
 /* org.apache.hadoop.hbase.util.HMerge */
8. protected HRegion merge(...)  \{
9.  HRegion r = HRegion.newHRegion(..., null);//\textcolor{red}{no FlushRequester}
10.  return r.openHRegion();
11.\} 
 \end{Verbatim}
 \vspace{-4mm}
 \\  \hline
 %\midrule
 \multicolumn{1}{c}{(C) A refactoring on the method and its caller} \\ \hline
 %\midrule
 Date: 8/31/2010

 Change Log: HBASE-2692 Unfinished renaming from CacheFlushListener to FlushRequester.
  \vspace{-2mm}
 \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
  /* org.apache.hadoop.hbase.regionserver.HRegion */
1. Context c = getContext();
2. public static HRegion newHRegion(..., 
3.     FlushRequester\textcolor{blue}{\underline{\underline{\bf flushRequester}}}) \{  
4.   if (\textcolor{blue}{\bf \uwave{flushRequester}} != null)  
5.     \textcolor{blue}{\bf \uwave{flushRequester}}.request(this);
6.   return c.newInstance(..., \textcolor{blue}{\bf \uwave{flushRequester}});  
7. \}
 \end{Verbatim} 
 %\vspace{-3mm}
 %\\  \hline
%\multicolumn{1}{c}{(C) Bug fix that is related to the inconsistent name} \\ \hline
%\midrule
Date: 9/8/2010

Change Log: FIX: Client can not send flush request.
\vspace{-2mm}
\begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
 /* org.apache.hadoop.hbase.util.HMerge */
 8. FlushRequester requester = new FlushRequester();
 9. protect HRegion merge(...)  \{
10. \textcolor{blue}{\bf HRegion r = HRegion.newHRegion(..., requester);  }
11.  return r.openHRegion();
12. \} 
 \end{Verbatim} 
 \vspace{-4mm}
 \\  \hline
 \end{tabular}
 \caption{Inconsistent parameter name detected with type-use pattern analysis}
 \label{fig:usageExample}
 \end{minipage}
 \end{figure}