### 2015-11-17 weeklyReport/table-RQ2.tex

 \begin{figure*}[!htb]
 \begin{minipage}{0.47\textwidth}
% \begin{table}[ht]
%\begin{center}
%\vspace{-5mm}
%\caption{Sampled inconsistent names and corresponding fixes.} 
%\label{tab:sample}
%\vspace{1mm}
\scriptsize{
\begin{tabular}{@{}l|p{3cm}p{3cm}}
Type&\multicolumn{2}{c}{(addition, fix revision, deletion)}\\  \midrule
%Pure&(7fac9,859056,---),&(a19676,c74517,---),\\
%&(0ccc3c,d5a9a3,---),&(542c7c,1d8a1d,ca1c68),\\
%&(b89bba,7e1fd3,---),&(37352e,939926c,---),\\
%&(a19676,a11042,---),&(74d393,ec7719,933a6d),\\
%&(8e4e6d,446a37,446a37),&(062620,ff2bcd,17dff6)\\   \midrule
%%(390f32,c9b028,---)
%Impure&(002dd69,e17cec,---),&(270277,d18d34e,ca1c68),\\ 
%&(dc372a,7c2d77,---),&(0201be,---,---),\\ 
%&(cd7157,---,---),&(cd7157,---,---), \\
%& (1daa81,---,48339b),&(55391f,---,ca1c68), \\
%&(0c5734,---,---),&(4ba168,---,---)\\ \midrule
%Object&(cd7157,fc7c8f,---),&(245343,ed3433,---),\\
%&(cd7157,1c309f,---),&(a6b82e,f51ed4,---),\\
%&(cf9afa, 36f7c4,36f7c4),&(4c792b,507d9d,507d9d),\\
%&(6aa468,54cccd,54cccd),&(ff7a62,d8b712,---), \\
%&(242f30,b3ba69,---),&(e1da90,f52130,f52130)\\\midrule
%(cd7157,d4fb88,---),(509bce,---,---)\\ 
%&(a6b82e,0a8ede,---), &(0930a8,c3fb42,---),\\
%&(cd7157,96247e,---), &(cd7157,47dfcf,---), \\
Pure&(\href{https://github.com/apache/hadoop/commit/7fac946#diff-d97658c87c6ac1d194da12d149834a25}{7fac94},\href{https://github.com/apache/hadoop/commit/8590564dc56195cb2caa245e3ee1c06eca3938d3#diff-52728ad908d378250fa082c8a30113af}{859056},---),&(\href{https://github.com/apache/hadoop/commit/a196766ea07775f18ded69bd9e8d239f8cfd3ccc#diff-9400d498326ceaca4b310cd3415277ef}{a19676},\href{https://github.com/apache/hadoop/commit/c74517#diff-649c4e46d99ae8011ff0fa0c118d9726}{c74517},---),\\
&(\href{https://github.com/apache/hadoop/commit/0ccc3c441a27646d761dfc913f750e2666f9ecfb#diff-1c2cc53548a274cf6eff37f3ded5dbbb}{0ccc3c},\href{https://github.com/apache/hadoop/commit/d5a9a3#diff-3f9d4f0c9157be654273622926ae04b9}{d5a9a3},---),&(\href{https://github.com/apache/hadoop/commit/542c7ccdce2680a4d179a8d948b8cc18221eb2bc#diff-eed8766da92e69dd6acd3fb9edc421f0}{542c7c},\href{https://github.com/apache/hadoop/commit/1d8a1d#diff-53b4c7b6b46b3778ebfa5ecdcb35521f}{1d8a1d},\href{https://github.com/apache/hadoop/commit/ca1c683b}{ca1c68}),\\
&(\href{https://github.com/apache/hadoop/commit/b89bba0acb#diff-cac0a345696e1521d21f8cbfd633f90c}{b89bba},\href{https://github.com/apache/hadoop/commit/7e1fd3#diff-336c3fa4df86402ec7b5339c1dbac724}{7e1fd3},---),&(\href{https://github.com/apache/hadoop/commit/37352ebe4dcbd526aafc343cccb350cd154e0315#diff-3de8a0425c49269f274f46497111dc80}{37352e},\href{https://github.com/apache/hadoop/commit/939926c#diff-3de8a0425c49269f274f46497111dc80}{939926},---),\\
&(\href{https://github.com/apache/hadoop/commit/a196766ea07775f18ded69bd9e8d239f8cfd3ccc#diff-6bfd89034be79e2f37aae9f211605fff}{a19676},\href{https://github.com/apache/hadoop/commit/a11042#diff-f471412017f4f6728224615c8362da5f}{a11042},---),&(\href{https://github.com/apache/hadoop/commit/74d393382341939}{74d393},\href{https://github.com/apache/hadoop/commit/ec7719#diff-2bd0b9b3e70406484c416d876341fa61}{ec7719},\href{https://github.com/apache/hadoop/commit/933a6d2c}{933a6d}),\\
&(\href{https://github.com/apache/accumulo/commit/8e4e6d6420e47f64b9fd54d9b6f5bc5b0b158b8b#diff-002fc1274cd98bf226ee8398551261a6}{8e4e6d},\href{https://github.com/apache/accumulo/commit/446a37a#diff-002fc1274cd98bf226ee8398551261a6}{446a37},\href{https://github.com/apache/accumulo/commit/446a37a#diff-002fc1274cd98bf226ee8398551261a6}{446a37}),&(\href{https://github.com/apache/hbase/commit/062660e3f48b#diff-76fffa96c664ed27a2f880fd02da3962}{062660},\href{https://github.com/apache/hbase/commit/ff2bcd#diff-53def65a3cf679a3f048ab71bcea9293}{ff2bcd},\href{https://github.com/apache/hbase/commit/17dff6#diff-53def65a3cf679a3f048ab71bcea9293}{17dff6})\\ \midrule
Impure&(\href{https://github.com/apache/hadoop/commit/002dd6968b89ded6a77858ccb50c9b2df074c226#diff-489fac1eb320e92dd65b62f00b67f71c}{002dd6},\href{https://github.com/apache/hadoop/commit/e17cec#diff-489fac1eb320e92dd65b62f00b67f71c}{e17cec},---),&(\href{https://github.com/apache/hadoop/commit/270277b9307c8de9df523e94bfc46df423e62b53#diff-7f17e003fb54ff97abf2007fbd6cdc61}{270277},\href{https://github.com/apache/hadoop/commit/d18d34e93#diff-7f17e003fb54ff97abf2007fbd6cdc61}{d18d34},\href{https://github.com/apache/hadoop/commit/ca1c68}{ca1c68}),\\
&(\href{https://github.com/apache/hadoop/blob/dc372af412a5923f04feaad8925a16d89180bc8d/src/java/org/apache/hadoop/mapred/TaskRunner.java}{dc372a},\href{https://github.com/apache/hadoop/commit/7c2d77#diff-c0169961c9fbca7c0a7b3afb2049665a}{7c2d77},---),&(\href{https://github.com/apache/hadoop/commit/0201be46c298e94176ec6297e9d9cdba3afc2bbd#diff-685393e5e55e645ac340ea5bfd3b1cf1}{0201be},---,---),\\
&(\href{https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ftp/FTPFileSystem.java}{cd7157},---,---),&(\href{https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ChecksumFileSystem.java}{cd7157},---,---),\\
&( \href{https://github.com/apache/hadoop/commit/1daa81#diff-13924c43e3ff5852da014b5005422666}{1daa81},---,---),&( \href{https://github.com/apache/hadoop/commit/55391f#diff-b7bcf8118ef56aa71a37a08b9479afd5}{55391f},---, \href{https://github.com/apache/hadoop/commit/ca1c68}{ca1c68}),\\
&(\href{https://github.com/apache/hadoop/commit/0c5734e4aca873f405fbf994e5fe7061e31731c8#diff-8da26f813ae9e87bbf0fb9abb349acc2}{0c5734},---,---),&(\href{https://github.com/apache/hadoop/commit/4ba168e4873c30574d18541b53b8821f62bd2589#diff-fa311023e9175bac1e31055102b7331f}{4ba168},---,---)\\ \midrule
Object&(\href{https://github.com/apache/hadoop/commit/5128a9a453d}{5128a9},\href{https://github.com/apache/hadoop/commit/fc7c8f9#diff-34d704a739ee5bdbee9bb5cd1426972b}{fc7c8f},---),&(\href{https://github.com/apache/hadoop/commit/2453431f642db9d38058b0227fd7636c162020e0#diff-72f498fd913780670fc13711bc524734}{245343},\href{https://github.com/apache/hadoop/commit/ed3433072#diff-72f498fd913780670fc13711bc524734}{ed3433},---),\\
&(\href{https://github.com/apache/hadoop/commit/cd7157784e5e5ddc4e77144d042e54dd0d04bac1#diff-16cbdb68a1f1e17dcc9f848ee6084e0c}{cd7157},\href{https://github.com/apache/hadoop/commit/1c309f763be3dd2e3d7d1616d2c960ff80cf9b03#diff-16cbdb68a1f1e17dcc9f848ee6084e0c}{1c309f},---),&(\href{https://github.com/apache/jackrabbit/commit/a6b82e}{a6b82e},\href{https://github.com/apache/jackrabbit/commit/f51ed4d48a#diff-9751fd2ad72989dbc11c7caca47e907c}{f51ed4},\href{https://github.com/apache/jackrabbit/commit/9cdb3c#diff-9751fd2ad72989dbc11c7caca47e907c}{9cdb3c}),\\
&(\href{https://github.com/apache/hadoop/commit/3512203f6c54033db1d52118edf595e5b5e1335e#diff-ad9e40bbf27e54506a64a57c691acc01}{351220},\href{https://github.com/apache/hadoop/commit/36f7c4360#diff-1c2cc53548a274cf6eff37f3ded5dbbb}{36f7c4},\href{https://github.com/apache/hadoop/commit/36f7c4360#diff-1c2cc53548a274cf6eff37f3ded5dbbb}{36f7c4}),&(\href{https://github.com/apache/commons-vfs/commit/c5451d949d67c415863565dcb56a6b2608a3aad0#diff-6ac9cecfbc1506096ed3445c48318795}{73e505},\href{https://github.com/apache/commons-vfs/commit/507d9d2061d4ea8c4ed4c26a5aad4f18aa493e67?diff=unified#diff-5a9a853727565aea8ec4f5670cafa120}{507d9d},\href{https://github.com/apache/commons-vfs/commit/507d9d2061d4ea8c4ed4c26a5aad4f18aa493e67?diff=unified#diff-5a9a853727565aea8ec4f5670cafa120}{507d9d}),\\
&(\href{https://github.com/apache/james/commit/6aa468ab069ce094027839b0282c47971598f28a#diff-53334ca0c55bca9fd6821a895f6ec46b}{6aa468},\href{https://github.com/apache/james/commit/54cccd3#diff-53334ca0c55bca9fd6821a895f6ec46b}{54cccd},\href{https://github.com/apache/james/commit/54cccd3#diff-53334ca0c55bca9fd6821a895f6ec46b}{54cccd}),&(\href{https://github.com/apache/jackrabbit/commit/ff7a62c0a8768714f8c36e3f8f07d2f017eebbfe#diff-7df848bb75ab7d3df5738968da840306}{ff7a62},\href{https://github.com/apache/jackrabbit/commit/d8b712#diff-7df848bb75ab7d3df5738968da840306}{d8b712},---),\\
&(\href{https://github.com/apache/hbase/commit/e1da900df#diff-b9d8ad451ab3fd5f5ff8593268230822}{e1da90},\href{https://github.com/apache/hbase/commit/f52130953245d93489a9c700c3b00eedbe70fdb9/src/main/java/org/apache/hadoop/hbase/regionserver/HRegion.java}{f52130},\href{https://github.com/apache/hbase/commit/f52130953245d93489a9c700c3b00eedbe70fdb9/src/main/java/org/apache/hadoop/hbase/regionserver/HRegion.java}{f52130}),&(\href{https://github.com/apache/hadoop/commit/a196766ea07775f18ded69bd9e8d239f8cfd3ccc#diff-9400d498326ceaca4b310cd3415277ef}{a19676},\href{https://github.com/apache/hadoop/commit/d4fb8821#diff-649c4e46d99ae8011ff0fa0c118d9726}{d4fb88},---)\\ \midrule

\end{tabular} \newline
--- represents that the inconsistent name was not fixed or deleted. 
 }
  \begin{itemize} 
\item \textbf{Pure term in impure methods.} In project \codefont{Apache accumulo}, a pure term \codefont{get} was used in an impure method \codefont{getEntries} that mutates its parameter \codefont{tablets} in an \codefont{if} condition. If this condition fails, this \codefont{getEntries} method will throw an exception to its caller, forcing the program to sleep for 3 seconds. This bug was fixed in revision \href{https://github.com/apache/accumulo/commit/446a37a#diff-002fc1274cd98bf226ee8398551261a6}{446a37} with a refactoring. 
\item \textbf{Impure term in a pure method.} As shown in Part (A) on the right, in project \codefont{Apache hadoop}, an impure term \codefont{prepare} was used in a pure method \codefont{prepare} which simply returns \codefont{true}.  When developer tried to add a new feature for \codefont{HADOOP-489}, she encountered a \codefont{NullPointerException} for the \codefont{prepare} method. The developer fixed this bug in revision \href{https://github.com/apache/hadoop/commit/7c2d77#diff-c0169961c9fbca7c0a7b3afb2049665a}{7c2d77}  by implementing another \codefont{init} method and invokes this new method in \codefont{prepare} method. This fix changes  the method \codefont{prepare} to impure method.
% In project \codefont{Apache hadoop}, an impure term \codefont{start} was used in a pure method \codefont{startSkipping}. The method \codefont{startSkipping} checks if the bad records can be skipped and returns the checking result. This method experiences  a bug fix which sets a maximum boundary for skipped records in revision \codefont{d18d34}.
\item \textbf{Inconsistent object names in similar type use patterns.} In project \codefont{Apache hbase}, there was a field named \codefont{flushListener} with type \codefont{FlushRequester}, while other object names with similar type use patterns used \codefont{*Requester} as a name. This inconsistent name was introduced in revision \href{https://github.com/apache/hbase/commit/e1da900df#diff-b9d8ad451ab3fd5f5ff8593268230822}{e1da90} when user renamed the class \codefont{CacheFlushListener} to \codefont{FlushRequester} yet forgot to rename the object name consistently. This inconsistent name was renamed as \codefont{flushRequester} in fix revision \href{https://github.com/apache/hbase/commit/f52130953245d93489a9c700c3b00eedbe70fdb9/src/main/java/org/apache/hadoop/hbase/regionserver/HRegion.java}{f52130}. 
 \end{itemize} 
% \vspace{-2mm}
% \end{center}
%  \vspace{-4mm}
%\end{table}
  \end{minipage}
  \hfill
 \begin{minipage}{0.47\textwidth}
 \scriptsize 
 \begin{tabular}{p{1\textwidth}}
  \hline 
  %https://github.com/apache/hadoop/commit/d18d34e9390cca94360675879fb29650e998d6e2/?diff=unified#diff-7f17e003fb54ff97abf2007fbd6cdc61
 % HADOOP-3829: If the number of skipped records in the neighborhood of a bad record are not 
%acceptable to the user, then narrow down the skipped range to the user acceptable value. 
   \multicolumn{1}{c}{(C) A impure term used in pure method, and fix on it.} \\ \hline
commit  \href{https://github.com/apache/hadoop/commit/7c2d77#diff-c0169961c9fbca7c0a7b3afb2049665a}{7c2d77}: HADOOP-489. Separate user logs from system logs.

\href{https://issues.apache.org/jira/browse/HADOOP-489}{HADOOP-489 issue comment}: ``my tasks fail with a NullPointer:..org.apache.hadoop.mapred.TaskRunner.prepare()..''
 \vspace*{-3mm}
 \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
 /* org.apache.hadoop.mapred.TaskRunner.java */
1.\color{red}{ -  public boolean prepare() throws IOException \{return true;\}}
2.\color{blue}{+  public boolean prepare() throws IOException \{}
3.\color{blue}{+    taskLogWriter.init();         }
4.\color{blue}{+    return true;}
5.\color{blue}{+  \}}
  \end{Verbatim}
  \vspace*{-4mm}
 \\  \hline
   \multicolumn{1}{c}{(B) A refactoring on the field with inconsistent name} \\ \hline
commit \href{https://github.com/apache/hbase/commit/f52130953245d93489a9c700c3b00eedbe70fdb9/src/main/java/org/apache/hadoop/hbase/regionserver/HRegion.java}{f52130}: HBASE-2692 Master rewrite and cleanup for 0.90.

SubTask issue description: \href{https://issues.apache.org/jira/browse/HBASE-2695}{HBASE-2695} Before doing the more significant changes to HMaster, it would benefit greatly from some cleanup, commenting, and a bit of refactoring.
 \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
 /* org.apache.hadoop.hbase.regionserver.HRegion.java */
1.\color{red}{-  final FlushRequester flushListener;}
2.\color{blue}{+  final FlushRequester flushRequester;}
3.  private void requestFlush() \{ 
4.\color{red}{-    if (this.flushListener == null) \{ }
5.\color{blue}{+    if (this.flushRequester == null) }
6.    \{ return; \}
7.\color{red}{-    this.flushListener.request(this);}
8.\color{blue}{+   this.flushRequester.requestFlush(this);}
9.\}
 \end{Verbatim}
  \vspace*{-4mm}
 \\  \hline 
  \multicolumn{1}{c}{(C) An impure term used in pure method and a bug fix on it} \\ \hline
commit  \href{https://github.com/apache/hadoop/commit/d18d34e93#diff-7f17e003fb54ff97abf2007fbd6cdc61}{d18d34}: HADOOP-3829. Narrown down skipped records based on user acceptable value.
 %\midrule
 \vspace*{-3mm}
 \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
 /*org.apache.hadoop.mapred.TaskInProgress.java: */
1. private boolean{\bf \hl{ startSkipping()}} \{
2.\color{red}{-  if(SkipBadRecords.getEnabled(conf) && }
3.\color{blue}{+  if(maxSkipRecords>0 && }
4.     numTaskFailures >= SkipBadRecords.
          getAttemptsToStartSkipping(conf)) \{
5.       return true;
6.     \} 
7.  return false;\}
\end{Verbatim}
 \vspace*{-4mm}
 \\  \hline
 %\midrule
% \multicolumn{1}{c}{(D) A impure term used in pure method, no bug fix on it.} \\ \hline
% \vspace*{-3mm}
% \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
%1. private int maximumFpRemove(int[] h) \{
%2.    int maxIndex = Integer.MIN_VALUE;
%3.    double maxValue = Double.MIN_VALUE;
%4.    for (int i = 0; i < nbHash; i++) \{
%5.      double fpWeight = getWeight(fpVector[h[i]]);
%6.      if (fpWeight > maxValue) \{
%7.        maxValue = fpWeight;
%8.        maxIndex = h[i];
%9.    \} \}
%10. return maxIndex; \}
% \end{Verbatim}
%  \vspace*{-4mm}
% \\  \hline 
  \end{tabular}
 \end{minipage}
  \begin{minipage}{1\textwidth}
%  \scriptsize{
Pure term used in impure methods: 
\begin{enumerate}
\item heartbeatCheck() in HeartbeatManager.java: addition \href{https://github.com/apache/hadoop/commit/7fac946#diff-d97658c87c6ac1d194da12d149834a25}{7fac94} fix \href{https://github.com/apache/hadoop/commit/8590564dc56195cb2caa245e3ee1c06eca3938d3#diff-52728ad908d378250fa082c8a30113af}{859056} no deletion.
\item receivePacket() in BlockReceiver.java: addition \href{https://github.com/apache/hadoop/commit/a196766ea07775f18ded69bd9e8d239f8cfd3ccc#diff-9400d498326ceaca4b310cd3415277ef}{a19676} fix \href{https://github.com/apache/hadoop/commit/c74517#diff-649c4e46d99ae8011ff0fa0c118d9726}{c74517} no deletion.
\item getFileChecksum() in DFSClient.java: addition \href{https://github.com/apache/hadoop/commit/0ccc3c441a27646d761dfc913f750e2666f9ecfb#diff-1c2cc53548a274cf6eff37f3ded5dbbb}{0ccc3c}  fix \href{https://github.com/apache/hadoop/commit/d5a9a3#diff-3f9d4f0c9157be654273622926ae04b9}{d5a9a3} no deletion.
\item obtainNewReduceTask() in JobInProgress.java: addition \href{https://github.com/apache/hadoop/commit/542c7ccdce2680a4d179a8d948b8cc18221eb2bc#diff-eed8766da92e69dd6acd3fb9edc421f0}{542c7c}  fix \href{https://github.com/apache/hadoop/commit/1d8a1d#diff-53b4c7b6b46b3778ebfa5ecdcb35521f}{1d8a1d} deletion \href{https://github.com/apache/hadoop/commit/ca1c683b}{ca1c68}.
\item fetchOutputs() in ReduceTask.java: addition \href{https://github.com/apache/hadoop/commit/b89bba0acb#diff-cac0a345696e1521d21f8cbfd633f90c}{b89bba} fix \href{https://github.com/apache/hadoop/commit/7e1fd3#diff-336c3fa4df86402ec7b5339c1dbac724}{7e1fd3} no deletion.  %%bug fix 01ad5a4d30f5d
\item checkLocalizedPath() in TestTrackerDistributedCacheManager.java: addition \href{https://github.com/apache/hadoop/commit/37352ebe4dcbd526aafc343cccb350cd154e0315#diff-3de8a0425c49269f274f46497111dc80}{37352e} fix \href{https://github.com/apache/hadoop/commit/939926c#diff-3de8a0425c49269f274f46497111dc80}{939926} no deletion.
\item verifyDir() in TestDFSUpgradeFromImage.java: addition \href{https://github.com/apache/hadoop/commit/a196766ea07775f18ded69bd9e8d239f8cfd3ccc#diff-6bfd89034be79e2f37aae9f211605fff}{a19676}  fix \href{https://github.com/apache/hadoop/commit/a11042#diff-f471412017f4f6728224615c8362da5f}{a11042} no deletion.
\item mapOutputLost() in TaskTracker.java:  addition \href{https://github.com/apache/hadoop/commit/74d393382341939}{74d393} fix  \href{https://github.com/apache/hadoop/commit/ec7719#diff-2bd0b9b3e70406484c416d876341fa61}{ec7719} deletion \href{https://github.com/apache/hadoop/commit/933a6d2c}{933a6d}
\item getEntries() in MetadataTable.java: addition \href{https://github.com/apache/accumulo/commit/8e4e6d6420e47f64b9fd54d9b6f5bc5b0b158b8b#diff-002fc1274cd98bf226ee8398551261a6}{8e4e6d}  fix and deletion \href{https://github.com/apache/accumulo/commit/446a37a#diff-002fc1274cd98bf226ee8398551261a6}{446a37}.
\item verifyRegionState() in Mocking.java: addition \href{https://github.com/apache/hbase/commit/062660e3f48b#diff-76fffa96c664ed27a2f880fd02da3962}{062660}  fix \href{https://github.com/apache/hbase/commit/ff2bcd#diff-53def65a3cf679a3f048ab71bcea9293}{ff2bcd} deletion \href{https://github.com/apache/hbase/commit/17dff6#diff-53def65a3cf679a3f048ab71bcea9293}{17dff6}
%\item getMetaTableRowResultAsSplitRegion() in Mocking.java: addition \href{}{} fix  \href{https://github.com/apache/hbase/commit/c9b028862a61ee2a96fdef6b55c36b03939ceada#diff-53def65a3cf679a3f048ab71bcea9293}{c9b028} 
%\item.  estimate() in PiEstimator.java:  fix \href{https://github.com/apache/hadoop/blob/509bcea8cf858a911a5285fe3acd0d000855e28a/src/examples/org/apache/hadoop/examples/PiEstimator.java}{509bce}
\end{enumerate}
%}
 \end{minipage}
  \end{figure*}[!htb]
  \begin{figure*}[!htb]
  \begin{minipage}{1\textwidth}
Impure term used in pure methods:
\begin{enumerate}
\item handleKind() in JobClient.java:  addition \href{https://github.com/apache/hadoop/commit/002dd6968b89ded6a77858ccb50c9b2df074c226#diff-489fac1eb320e92dd65b62f00b67f71c}{002dd6} fix  \href{https://github.com/apache/hadoop/commit/e17cec#diff-489fac1eb320e92dd65b62f00b67f71c}{e17cec} no deletion.
\item  startSkipping() in TaskInProgress.java addition \href{https://github.com/apache/hadoop/commit/270277b9307c8de9df523e94bfc46df423e62b53#diff-7f17e003fb54ff97abf2007fbd6cdc61}{270277} fix  \href{https://github.com/apache/hadoop/commit/d18d34e93#diff-7f17e003fb54ff97abf2007fbd6cdc61}{d18d34} deletion \href{https://github.com/apache/hadoop/commit/ca1c68}{ca1c68}.
\item  prepare() in TaskRunner.java  addition \href{https://github.com/apache/hadoop/blob/dc372af412a5923f04feaad8925a16d89180bc8d/src/java/org/apache/hadoop/mapred/TaskRunner.java}{dc372a} fix (this inconsistent name is fixed by implementation change)   \href{https://github.com/apache/hadoop/commit/7c2d77#diff-c0169961c9fbca7c0a7b3afb2049665a}{7c2d77} no deletion
\item  open() in HarFileSystem.java addition \href{https://github.com/apache/hadoop/commit/0201be46c298e94176ec6297e9d9cdba3afc2bbd#diff-685393e5e55e645ac340ea5bfd3b1cf1}{0201be}  no fix no deletion.
% \href{https://github.com/apache/hadoop/commit/5128a9a453d64bfe1ed978cf9ffed27985eeef36#diff-c889130fbb61593f9344d4455243a5ba}{5128a9a} fix \href{https://github.com/apache/hadoop/blob/0201be46c298e94176ec6297e9d9cdba3afc2bbd/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/HarFileSystem.java}{0201be}
\item  setWorkingDirectory() in FTPFileSystem.java. addition \href{https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ftp/FTPFileSystem.java}{cd7157}  no fix no deletion.
%fix \href{https://github.com/apache/hadoop/commit/e09ea0c06ee1caa5a9ebae0a8f0273dfe04d05e5#diff-ff793a007e42565e9d1844ce96940948}{e09ea0c0}.  latest version \href{https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ftp/FTPFileSystem.java}{cd7157}
\item  startLocalOutput() in ChecksumFileSystem.java addition \href{https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ChecksumFileSystem.java}{cd7157} no fix no deletion
\item  invertedMinMax() in PoolManager.java addition  \href{https://github.com/apache/hadoop/commit/1daa81#diff-13924c43e3ff5852da014b5005422666}{1daa81} no fix no deletion
\item  ranAll() in JvmManager.java  addition \href{https://github.com/apache/hadoop/commit/55391f#diff-b7bcf8118ef56aa71a37a08b9479afd5}{55391f} no fix deletion \href{https://github.com/apache/hadoop/commit/ca1c68}{ca1c68}
\item  abort() in UserGroupInformation.java addition \href{https://github.com/apache/hadoop/commit/0c5734e4aca873f405fbf994e5fe7061e31731c8#diff-8da26f813ae9e87bbf0fb9abb349acc2}{0c5734}  no fix no deletion
\item  maximumFpRemove() in RetouchedBloomFilter.java  addition \href{https://github.com/apache/hadoop/commit/4ba168e4873c30574d18541b53b8821f62bd2589#diff-fa311023e9175bac1e31055102b7331f}{4ba168} no fix no deletion
\end{enumerate}
Object
\begin{enumerate}
\item  SequenceFile.Writer data in MapFile.java addition \href{https://github.com/apache/hadoop/commit/5128a9a453d}{5128a9}   fix \href{https://github.com/apache/hadoop/commit/fc7c8f9#diff-34d704a739ee5bdbee9bb5cd1426972b}{fc7c8f} no deletion.
\item Process sim in PipeMapRed.java  addition \href{https://github.com/apache/hadoop/commit/2453431f642db9d38058b0227fd7636c162020e0#diff-72f498fd913780670fc13711bc524734}{245343}   fix  \href{https://github.com/apache/hadoop/commit/ed3433072#diff-72f498fd913780670fc13711bc524734}{ed3433} no deletion
\item  FSDataInputStream sums in ChecksumFileSystem.java  addition \href{https://github.com/apache/hadoop/commit/cd7157784e5e5ddc4e77144d042e54dd0d04bac1#diff-16cbdb68a1f1e17dcc9f848ee6084e0c}{cd7157}  fix \href{https://github.com/apache/hadoop/commit/1c309f763be3dd2e3d7d1616d2c960ff80cf9b03#diff-16cbdb68a1f1e17dcc9f848ee6084e0c}{1c309f}  no deletion
\item  FileSystem repStore in RepositoryImpl.java addition \href{https://github.com/apache/jackrabbit/commit/a6b82e}{a6b82e}  fix  \href{https://github.com/apache/jackrabbit/commit/f51ed4d48a#diff-9751fd2ad72989dbc11c7caca47e907c}{f51ed4} deletion in type-use pattern \href{https://github.com/apache/jackrabbit/commit/f51ed4#diff-9751fd2ad72989dbc11c7caca47e907c}{f51ed4} complete deletion \href{https://github.com/apache/jackrabbit/commit/9cdb3c#diff-9751fd2ad72989dbc11c7caca47e907c}{9cdb3c}
\item SortedMap<String, OutputStream> pendingCreates in DFSClient.java bug addition \href{https://github.com/apache/hadoop/commit/3512203f6c54033db1d52118edf595e5b5e1335e#diff-ad9e40bbf27e54506a64a57c691acc01}{351220}   fix and deletion \href{https://github.com/apache/hadoop/commit/36f7c4360#diff-1c2cc53548a274cf6eff37f3ded5dbbb}{36f7c4} 
\item  FileObject webdavTmp in WebdavFileObject.java addition  \href{https://github.com/apache/commons-vfs/commit/c5451d949d67c415863565dcb56a6b2608a3aad0#diff-6ac9cecfbc1506096ed3445c48318795}{73e505}  fix and deletion \href{https://github.com/apache/commons-vfs/commit/507d9d2061d4ea8c4ed4c26a5aad4f18aa493e67?diff=unified#diff-5a9a853727565aea8ec4f5670cafa120}{507d9d}  %deletion \href{https://github.com/apache/commons-vfs/commit/507d9d2061d4#diff-8adfff94d4a9ad884d8367d5d25e8864}{507d9d}
\item  SpoolRepository outgoing in RemoteDelivery.java addition \href{https://github.com/apache/james/commit/6aa468ab069ce094027839b0282c47971598f28a#diff-53334ca0c55bca9fd6821a895f6ec46b}{6aa468}  bug fix and deletion \href{https://github.com/apache/james/commit/54cccd3#diff-53334ca0c55bca9fd6821a895f6ec46b}{54cccd} 
\item  PersistenceManager target in PersistenceCopier.java addition \href{https://github.com/apache/jackrabbit/commit/ff7a62c0a8768714f8c36e3f8f07d2f017eebbfe#diff-7df848bb75ab7d3df5738968da840306}{ff7a62} fix  \href{https://github.com/apache/jackrabbit/commit/d8b712#diff-7df848bb75ab7d3df5738968da840306}{d8b712}  no deletion
\item  FlushRequester flushListener in HRegion.java addition \href{https://github.com/apache/hbase/commit/e1da900df#diff-b9d8ad451ab3fd5f5ff8593268230822}{e1da90}  fix and deletion \href{https://github.com/apache/hbase/commit/f52130953245d93489a9c700c3b00eedbe70fdb9/src/main/java/org/apache/hadoop/hbase/regionserver/HRegion.java}{f52130} 
\item  Daemon responder in BlockReceiver.java addition \href{https://github.com/apache/hadoop/commit/a196766ea07775f18ded69bd9e8d239f8cfd3ccc#diff-9400d498326ceaca4b310cd3415277ef}{a19676}  fix \href{https://github.com/apache/hadoop/commit/d4fb8821#diff-649c4e46d99ae8011ff0fa0c118d9726}{d4fb88}  no deletion 
\item  File repositoryHome in FileJournalTest.java addition \href{https://github.com/apache/jackrabbit/commit/242f30ca639d0830e5ffe8f38c310453210934dd#diff-0ec8ddbe8d780a64761634b465905590}{242f30}  fix \href{https://github.com/apache/jackrabbit/commit/b3ba69bce1#diff-0ec8ddbe8d780a64761634b465905590}{b3ba69}    no deletion
  \end{enumerate}
%}
  \end{minipage}
 % \caption{Inconsistent method name detected with purity analysis and type-use pattern}
% \label{fig:purityExample}
 \end{figure*}

% As a result, developers did not notice that \codefont{flashListener} was a \codefont{FlashRequester} and forgot to pass it to \codefont{request} method, {causing a bug where a user cannot send request via Requester}. This bug was fixed in revision \codefont{d4fb88}.

%%Pure: 
%1. heartbeatCheck() in HeartbeatManager.java: 859056 https://github.com/apache/hadoop/commit/8590564dc56195cb2caa245e3ee1c06eca3938d3#diff-52728ad908d378250fa082c8a30113af. 
%2. receivePacket() in BlockReceiver.java: c74517 https://github.com/apache/hadoop/commit/c74517#diff-649c4e46d99ae8011ff0fa0c118d9726
%3. getFileChecksum() in DFSClient.java: d5a9a3 https://github.com/apache/hadoop/commit/d5a9a3#diff-3f9d4f0c9157be654273622926ae04b9
%4. obtainNewReduceTask() in JobInProgress.java: 1d8a1d https://github.com/apache/hadoop/commit/1d8a1d#diff-53b4c7b6b46b3778ebfa5ecdcb35521f
%5. fetchOutputs() in ReduceTask.java: 7e1fd3 https://github.com/apache/hadoop/commit/7e1fd3#diff-336c3fa4df86402ec7b5339c1dbac724
%6. checkLocalizedPath() in TestTrackerDistributedCacheManager.java: 939926c https://github.com/apache/hadoop/commit/939926c#diff-3de8a0425c49269f274f46497111dc80
%7. verifyDir() in TestDFSUpgradeFromImage.java: a11042 https://github.com/apache/hadoop/commit/a11042#diff-f471412017f4f6728224615c8362da5f 
%8. mapOutputLost() in TaskTracker.java: ec7719 https://github.com/apache/hadoop/commit/ec7719#diff-2bd0b9b3e70406484c416d876341fa61
%9. getEntries() in MetadataTable.java: bug fix and deletion https://github.com/apache/accumulo/commit/446a37a#diff-002fc1274cd98bf226ee8398551261a6 additiond in https://github.com/apache/accumulo/commit/8e4e6d6420e47f64b9fd54d9b6f5bc5b0b158b8b#diff-002fc1274cd98bf226ee8398551261a6
%10. getMetaTableRowResultAsSplitRegion() in ZKAssign.java: c9b028  https://github.com/apache/hbase/commit/c9b028862a61ee2a96fdef6b55c36b03939ceada#diff-53def65a3cf679a3f048ab71bcea9293

%9.  estimate() in PiEstimator.java: 509bce https://github.com/apache/hadoop/blob/509bcea8cf858a911a5285fe3acd0d000855e28a/src/examples/org/apache/hadoop/examples/PiEstimator.java

%%Impure:
%1. handleKind() in JobClient.java e17cec: https://github.com/apache/hadoop/commit/e17cec#diff-489fac1eb320e92dd65b62f00b67f71c
%2. startSkipping() in TaskInProgress.java d18d34e93: https://github.com/apache/hadoop/commit/d18d34e93#diff-7f17e003fb54ff97abf2007fbd6cdc61
%3. prepare() in TaskRunner.java 7c2d77: https://github.com/apache/hadoop/commit/7c2d77#diff-c0169961c9fbca7c0a7b3afb2049665a
%4. open() in HarFileSystem.java no fix. https://github.com/apache/hadoop/blob/0201be46c298e94176ec6297e9d9cdba3afc2bbd/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/HarFileSystem.java
%5. setWorkingDirectory() in FTPFileSystem.java. no fix. cd7157 latest version https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ftp/FTPFileSystem.java
%6. startLocalOutput() in ChecksumFileSystem.java cd71577 no fix latest version https://github.com/apache/hadoop/blob/3b85bd7c3391719624a261f061a67cc4356be477/hadoop-common-project/hadoop-common/src/main/java/org/apache/hadoop/fs/ChecksumFileSystem.java
%7. invertedMinMax() in PoolManager.java no fix. 1daa81. https://github.com/apache/hadoop/commit/1daa81#diff-13924c43e3ff5852da014b5005422666
%8. ranAll() in JvmManager.java 55391f:  https://github.com/apache/hadoop/commit/55391f#diff-b7bcf8118ef56aa71a37a08b9479afd5
%9. abort() in UserGroupInformation.java 0c5734e https://github.com/apache/hadoop/commit/0c5734e4aca873f405fbf994e5fe7061e31731c8#diff-8da26f813ae9e87bbf0fb9abb349acc2
%10. maximumFpRemove() in RetouchedBloomFilter.java 4ba168e48 no fix. https://github.com/apache/hadoop/commit/4ba168e4873c30574d18541b53b8821f62bd2589#diff-fa311023e9175bac1e31055102b7331f
%%Object
%1. SequenceFile.Writer data in MapFile.java fc7c8f9 bug fix https://github.com/apache/hadoop/commit/fc7c8f9#diff-34d704a739ee5bdbee9bb5cd1426972b
%2. Process sim in PipeMapRed.java bug fix ed3433072 https://github.com/apache/hadoop/commit/ed3433072#diff-72f498fd913780670fc13711bc524734 additiond in https://github.com/apache/hadoop/commit/2453431f642db9d38058b0227fd7636c162020e0#diff-72f498fd913780670fc13711bc524734
%3. FSDataInputStream sums in ChecksumFileSystem.java 1c309f bug fix https://github.com/apache/hadoop/commit/1c309f763be3dd2e3d7d1616d2c960ff80cf9b03#diff-16cbdb68a1f1e17dcc9f848ee6084e0c additiond in https://github.com/apache/hadoop/commit/cd7157784e5e5ddc4e77144d042e54dd0d04bac1#diff-16cbdb68a1f1e17dcc9f848ee6084e0c
%4. FileSystem repStore in RepositoryImpl.java bug fix  https://github.com/apache/jackrabbit/commit/f51ed4d48a#diff-9751fd2ad72989dbc11c7caca47e907c
%5. SortedMap<String, OutputStream> pendingCreates in DFSClient.java bug fix and deletion https://github.com/apache/hadoop/commit/36f7c4360#diff-1c2cc53548a274cf6eff37f3ded5dbbb
%6. FileObject webdavTmp in WebdavFileObject.java bug fix and deletion https://github.com/apache/commons-vfs/commit/507d9d2061d4ea8c4ed4c26a5aad4f18aa493e67?diff=unified#diff-5a9a853727565aea8ec4f5670cafa120
%7. SpoolRepository outgoing in RemoteDelivery.java bug fix and deletion https://github.com/apache/james/commit/54cccd3#diff-53334ca0c55bca9fd6821a895f6ec46b additiond in https://github.com/apache/james/commit/6aa468ab069ce094027839b0282c47971598f28a#diff-53334ca0c55bca9fd6821a895f6ec46b
%8. PersistenceManager target in PersistenceCopier.java  bug fix  https://github.com/apache/jackrabbit/commit/d8b712#diff-7df848bb75ab7d3df5738968da840306 additiond in https://github.com/apache/jackrabbit/commit/ff7a62c0a8768714f8c36e3f8f07d2f017eebbfe#diff-7df848bb75ab7d3df5738968da840306
%9. FlushRequester flushListener in HRegion.java bug fix and deletion https://github.com/apache/hbase/commit/f52130953245d93489a9c700c3b00eedbe70fdb9/src/main/java/org/apache/hadoop/hbase/regionserver/HRegion.java additiond in https://github.com/apache/hbase/commit/e1da900df#diff-b9d8ad451ab3fd5f5ff8593268230822 
%10. Daemon responder in BlockReceiver.java bug fix https://github.com/apache/hadoop/commit/d4fb8821#diff-649c4e46d99ae8011ff0fa0c118d9726 
% File repositoryHome in FileJournalTest.java bug fix https://github.com/apache/jackrabbit/commit/b3ba69bce1#diff-0ec8ddbe8d780a64761634b465905590 additiond in https://github.com/apache/jackrabbit/commit/242f30ca639d0830e5ffe8f38c310453210934dd#diff-0ec8ddbe8d780a64761634b465905590