### 2017-10-01 sketchAPI/paper/algorithm.tex
rename ase17

\begin{algorithm}[t!]
\small
\SetKwInOut{Input}{Input}
\SetKwInOut{Output}{Output} 
\SetKwProg{Prog}{Title}{ is}{end}
\SetKwProg{Fn}{Function}{ is}{end}
\SetKw{New}{new}
\SetKw{Throw}{throw}
\SetKw{Continue}{continue}
\SetKwIF{Trying}{Catch}{Catching}{try}{}{catch}{}{catch}{}%
%\SetKwIF{Try}{Catch}{try}{catch}{}%
\SetKwRepeat{Do}{do}{while}{}{}%
\Input {Maximum Length of the sequence $len$, Input arguments $args$}
\Output{ API Sequence candidate $sequence$ and the Return Value $returnVal$ }
\SetKwProg{Fn}{Function}{ is}{end}
\Fn{ genAPISequence (hole)  }{
rule $\leftarrow$ genConstraints(hole.len).choose()\label{line:chooseRule}\;
sequence  $\leftarrow  \varnothing$, args $\leftarrow$ hole.args\;
\For {int i=0; i$<$rule.stmt; i++}{
call $\leftarrow $ genCall(rule, hole, i)\label{line:genCall}\;
\Trying{}{ 
res $\leftarrow$ invoke(call, args)\;  
args $\leftarrow$  args $\cup$ res\;  }
\Catch{Exception e}{\Throw BacktrackException\label{line:genBacktrack};}
sequence $\leftarrow$ sequence  $\cup$  call\;
\Return $\langle$sequence, getReturnVal(rule, args)$\rangle$;
}}
\Fn{genCall(rule, hole, i)}{
\tcp{Generate API sketch}\
\If{hole.mtdCandidates ==  $\varnothing$}{
methods $\leftarrow$ fetchMethods(hole.sourceTypes)\label{line:fetchMethod}\;
\ForEach{ m $\in$ methods}{\If{violateConstraint(hole, rule, i)}{\Continue;}
%\If{cannotFitRequiredArgs(m, hole.args)}{\Continue;}
hole.mtdCandidates $\leftarrow $ hole.mtdCandidates $\cup$ m;
} 
}
\tcp{Complete expressions holes in methods}\
method $\leftarrow $ hole.mtdCandidates.choose()\;
call $\leftarrow$fillMethod(method, hole, i, rule).choose()~\label{line:fillExpr}\;
\Return call;
}
\Fn{ execute (hole)  }{
\If{hole.sequence==null}{\label{alg:candNull}
hole.len $\leftarrow$ hole.len $== -1 ?$ b: hole.len\label{line:holeLen}\;
$\langle$sequence, returnVal$\rangle  \leftarrow$ genSequence(hole)\;
hole.sequence $\leftarrow$ sequence\;
\Return returnVal;
}\Else{
\Return invokeAndReturn(hole);
}
}
\Fn{ genConstraints ($len$) \label{line:genrule} }{
\tcp{\#stmt, \#return, \#consume, \#repetition}
List$<$int[]$>$ rules $\leftarrow   \varnothing$\;
\For{int stmt=1; stmt$<$len; stmt++}{
\For{int rtn=stmt; rtn$>$0; rtn$--$}{\For{int vc=1; vc$<$stmt; vc++}{
\For{int rep=1; rep$<$stmt; rep++}{
rules $\leftarrow $ rules $\cup$ \{stmt, rtn, vc, rep\};
}}}}
sort(rules)\;
\Return rules;
}
\caption{API Candidate Generation}
\label{alm:genCandidate}
\end{algorithm}