### 2017-10-01 sketchAPI/paper/example_background.tex

\begin{figure}[htb]
\footnotesize
      \begin{minipage}{0.5\textwidth}
      \begin{tabular}{@{}p{0.9\textwidth}}
 \\ \hline
  \multicolumn{1}{c}{(A) A method that contains an API sequence under synthesis } \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
//Compute the ith  eigenvalue of a matrix
1.Vector2D eigenvalue(RealMatrix arg0, int arg1) \{
2. //Synthesize a sequence of APIs 
   //for desired functionality
3.\}
 \end{Verbatim}  
  \\ \hline
  \multicolumn{1}{c}{(B) A given test case that specifies desired behavior  } \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
1.public static boolean test0() throws Throwable \{
2. double[][] mat = new double[][]\{\{0,-20\},\{10,10\}\};
3. RealMatrix matrix = new Array2DRowRealMatrix(mat);
4. Vector2D result = eigenvalue(matrix, 0);
5. Vector2D target = new Vector2D(5, 5*Math.sqrt(7));
6. return Math.abs(result.getX()-target.getX())<1e-6 
 \&\& Math.abs(result.getY()-target.getY())<1e-6;
7.\}
 \end{Verbatim}  
  \\ \hline
  \multicolumn{1}{c}{(C) Input Sketch for \tool  Written by Users} \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
1.Vector2D eigenvalue(RealMatrix arg0, int arg1) \{
2. TedSketch.INVOKE(0) /* Hole ID*/
3. .addParameter(RealMatrix.class, arg0)
4. .addParameter(int.class, arg1)
5. .setReturnType(Vector2D.class)
6. .addPackage("org.apache.commons.math3.linear")
7. .invoke();
8.\}
 \end{Verbatim}
        \\ \hline
  \multicolumn{1}{c}{(D) Input Configuration JSON File for \spt Written by Users} \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
1. \{ 
2.  "methodName": "eigenvalue",
3.  "srcTypes": [	"RealMatrix", "int"],
4.  "paramNames": ["arg0", "arg1"],
5.  "tgtType": "Vector2D",
6.  "packages": ["org.apache.commons.math3.linear"],
7.  "testPath": "TestSource.java"
8.\}
 \end{Verbatim}
   \\ \hline
  \multicolumn{1}{c}{(E) A solution generated by \tool based on the test suite  } \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
1.TedSketch.INVOKE(0)
2. EigenDecomposition res0 = new 
    EigenDecomposition(arg0);
3. double res1 = res0.getImagEigenvalue(arg1);
4. double res2 = res0.getRealEigenvalue(arg1);
5. Vector2D res3 = new Vector2D(res2, res1);
6.return res3;
 \end{Verbatim}  
 \\  \hline
 \end{tabular}
  \end{minipage}
   \caption{An synthesis example of straight-line API sequence}
 \label{fig:background}
 \end{figure}