### 2017-10-01 sketchAPI/paper/example_partition.tex

\begin{figure}[t!]
\footnotesize
      \begin{minipage}{0.5\textwidth}
      \begin{tabular}{@{}p{0.9\textwidth}}

   \\ \hline
  \multicolumn{1}{c}{(A) A code skeleton under synthesis as the input of \spt} \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
1.String classify(int a, int b, int c) \{ 
2. if(condition1(a, b, c)) \{
3.  return body1();
4. \} else if(condition2(a, b, c)) \{ 
5.  return body2();
6. \} else \{
7.  return body3();
8. \} \}
 \end{Verbatim}
   \\ \hline
  \multicolumn{1}{c}{(B) Sample test cases that define correct behavior} \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
public static boolean test1() throws Throwable \{
 return classify(78, 79, 80).equals("acute");
\}
public static boolean test2() throws Throwable \{
 return classify(12, 13, 5).equals("right");
\}
public static boolean test3() throws Throwable \{
 return classify(10, 15, 10).equals("obtuse");    
\}
 \end{Verbatim}  
      \\ \hline
  \multicolumn{1}{c}{(B) A solution from extended \spt  } \\ \hline
  \begin{Verbatim}[commandchars=\\\{\}, tabsize=2]
1. body3(a,b,c): Library.obtuse(a,b,c);
2. body2(a,b,c): Library.right(a,b,c);
3. body1(a,b,c): Library.acute(a,b,c);
4. condition2(a,b,c): Library.isRight(a,b,c);
5. condition1(a,b,c): Library.isAcute(a,b,c);
 \end{Verbatim}
 \\  \hline
 \end{tabular}
  \end{minipage}
   \caption{Example of \spt extension to synthesize non-straight-line APIs   }
 \label{fig:partition}

 \end{figure}