### 2017-10-01 sketchAPI/paper/table_line.tex

\begin{table*}[!htb]
\footnotesize
\begin{center}
\caption{ Straight-Line Subjects}
\label{table:line}
\begin{tabular}{@{}c|l|r|rr|rr|rr|rrrrrrrrrrrr@{}}

\multirow{3}{*}{ID}&	\multirow{3}{*}{Description}&\multirow{3}{*}{\#Tests}&\multicolumn{2}{c|}{\spt}&\multicolumn{6}{c}{\tool} \\
&&&\multirow{2}{*}{\#API}& \multirow{2}{*}{Space} &\multirow{2}{*}{\#API}&\multirow{2}{*}{Space}&\multicolumn{2}{c|}{With Ranking}&\multicolumn{2}{c}{Without Ranking}\\ 
&&&&&& &Time(s)&\#Run&Time(s)&\#Run\\ \midrule
1&Compute the pseudo-inverse of a matrix&1&3&41&3&1.7b&6&1.8k&12&5.3k\\
2&Compute the inner product between two vectors&1&3&24&2&258.6m&2&124&3&134\\
3&Determine the roots of a polynomial equation&1&3&35&3&23.1t&109&9.0k&380&69.3k\\
4&Compute the singular value decomposition of a matrix&1&3&24&3&45.5b&3&286&8&993\\
5&Invert a square matrix&1&3&28&2&12.7m&2&135&3&217\\
6&Solve a system of linear equations&1&6&115&-& $10^{22}$&-&146.7k&-&959.7k\\
7&Compute the outer product between two vectors&1&4&37&2&258.6m&2&117&3&127\\
8&Predict a value from a sample by linear regression&2&3&244&3&18.1b&517&3.7k&4&3.8k\\
9&Compute the ith eigenvalue of a matrix&2&-&$\perp$&4&$10^{15}$&62&211.2k&339&850.3k\\
10&Scale a rectangle by a given ratio&1&4&36&3&502.0b&314&32.0m&710&18.3m\\
11&Shear a rectangle and get its bounds&1&4&36&3&502.0b&308&31.9m&133&13.1m\\
12&Rotate a rectangle about the origin by quadrants&1&4&17&3&32.2b&2&907&3&2.6k\\
13&Rotate 2-D shape by the specified angle about a point&2&4&33&-&$10^{15}$&-&19.7m&-&18.2m\\
14&Perform a translation on a rectangle&1&4&26&3&502.0b&308&32.2m&-&55.0m\\
15&Intersect a rectangle and an ellipse&1&3&11&2&10.1m&2&40&1&3\\
16&Compute number of days since a date&2&3&22&3&45.8b&37&80.8k&70&111.6k\\
17&Subtract two dates considering timezone&3&4&408&3&2.3t&56&150.4k&122&214.2k\\
18&Determine if a year is a leap year&3&4&68&3&45.8b&24&102.3k&3&1.2k\\
19&Return the day of a date string&2&3&11&3&2.5t&42&252.9k&43&57.7k\\
20&Find the number of days of a month in a date string&2&4&83&-& $10^{16}$&-&38.1m&-&5.6m\\
21&Find the day of the week of a date string&2&4&44&-& $10^{16}$&-&19.4m&-&5.4m\\
22&Compute age given date of birth&2&3&30&3&45.8b&63&158.6k&86&174.1k\\
23&Compute the offset for a specified line in a document&1&3&19&3&1.7t&5&1.2k&11&4.0k\\
24&Get a paragraph element given its offset in the a document&1&3&24&3&1.7t&4&1k&11&3.8k\\
25&Obtain the title of a webpage specified by a URL&1&3&110&3&1.1b&35&10.7k&35&19.2k\\
26&Return doctype of XML document generated by string&1&6&21&-& $10^{19}$&-&5.4m&-&5.6m\\
27&Generate an XML element from a string&1&6&24&-& $10^{19}$&-&3.8m&-&4.8m\\
28&Read XML document from a file&1&3&14&3&5.5b&4&3.9k&5&2.8k\\
29&Generate an XML from file and query it using XPath&1&6&78&-& $10^{26}$&-&3.2m&-&3.3m\\
30&Get the value of root attribute from a XML file&1&5&17&-&$10^{18}$&-&13.0m&-&13.8m\\ \midrule
31&Check if a point is inside a rectangle&8&5&66&1&3.1k&0.5&1&0.4&1\\
32&Check if a line segment intersects a rectangle.&8&-&$\perp$&2&63.1m&1&224&2&265\\
33&Compute number of minutes between two time&8&-&$\perp$&2&175.5m&1&183&1&350\\
34&Get number of seconds since the midnight of sometime&8&2&15&1&3.5k&1&2&0.7&2\\
35&Compute the transpose of a matrix&8&3&12&3&1.7b&1&88&1&401\\
36&Compute the sum of two matrices&8&4&14&4&911.1t&99&3.8k&1.4k&56.5k\\
37&Compute exclusive or between an area and a rectangle&2&-&-&2&10.1m&3&135&3&136\\
38&Create an element with given name&3&4&18&4&3.3t&5&6.5k&1&779\\ \midrule
39&Calculate absolute value of an integer&8&1&15&1&2.4k&0.5&5&0.1&10\\
40&Increment an integer and return its old value&8&4&5&3&1.9k&0.4&60&0.9&28\\
41&Increment an integer by 2 and return its old value&8&-&-&5&271.5k&7&115.9k&0.5&110\\
42&Get the class name of an object&8&2&30&2&5.9m&0.5&25&0.6&28\\
43&Get the first value of an integer array&8&2&4&2&210&1&14&0.8&8\\
44&Calculate minimum value between two integers&8&1&14&1&7.1k&0.5&49&0.5&49\\
45&Calculate minimum value between three integers&8&2&25&2&179.8m&1.1&3.8k&1&4.3k\\
46&Given an array set the last entry the value of first entry&8&4&154&5&579.2k&1&3.1k&2&16.4k\\
47&Sort an integer array&8&1&14&1&7.6k&0.8&1&1&1\\

  \end{tabular}

  \vspace{1mm}
 $\perp$ represents out of memory. - represents time out after 30 minutes. The first 30 subjects are from the evaluation benchmark of \spt, 31-38 are 8 subjects derived from the open source projects used in \spt evaluation, and the rest are small synthesis tasks such as absolute value calculation.
\end{center}
\end{table*}