### 2017-10-01 sketchAPI/paper/table_sub.tex

\begin{table*}
\footnotesize
\begin{center}
\caption{ Non-Straight-Line Subjects}
\label{table:branch}
\begin{tabular}{@{}c|c|l|rrrrrrrr@{}}
Project&ID&Description&\#Cond 	&\#Stmt&Space&\#Tests&\#Run&Time(s)\\ \midrule
Chart&1&\codefont{createAndAddEntity()}:  Created an entity for the axis&3if&6&5.5$\times10^{21}$&1&279&9\\
90k; 2.2k&2&\codefont{addBaseTimelineException()}:  Adds a segment &1wh&1&2.4k&2&2&11\\ \midrule
Math&3&\codefont{ebeDivide()}:  Element-by-element division in a vector&1wh&4&9.4b&1&7.8k&50\\
Loc: 85k&4&\codefont{add()}: Optimized method to add two RealVectors&2if, 1wh&6&8.4$\times10^{19}$&1&170.5k&51\\
\#Tests: 3.6k&5&\codefont{setEntry()}: Set entry in specified row and column&3if&2&9.3$\times10^{14}$&1&529&33\\  \midrule
&6&\codefont{visit()}:   sets properties as ineligible&1if&2&9k&57&1&3\\
Closure&7&\codefont{applyCollapses()}:  Collapse  variable declarations &1if, 1wh&2&12.5b&6&125.0k&220\\
Loc: 90k&8&\codefont{remove()}: Remove this node&1if&3&1.4$\times10^{11}$&10&172.4k&333\\
\#Tests: 7.8k&9&\codefont{flattenReferences()}:  Flattens to collapsible properties&2if&2&7.0m&23&9&6\\
&10&\codefont{removeVar()}:  remove var if it has been coalesced&3if&3&1.8$\times10^{14}$&28&195.1k&2.3k\\
  \end{tabular}
  \vspace{1mm}

  *if represents the total number of synthesized APIs in if-conditions, and *wh represents the total number of generated methods in while loops.
\end{center}
\end{table*}